PineAPPL: combining EW and QCD corrections\\ for fast evaluation of LHC processes

### 2020-09-03 paper/sec-introduction.tex
Addressed Valerio's complaint

In this paper we address the first of these two problems. Specifically, 
we develop \textsc{PineAPPL}, a library that allows any user to generate
fast-interpolation grids, accurate to any fixed order in the QCD and
EW couplings. The library supports variations of the factorisation
and renormalisation scales, and can be extended to include
resummation, and matching with a photon- and/or parton-shower.
The grids in the new \textsc{PineAPPL} format complement those
(accurate to fixed order in the strong coupling only) that can be generated
in the \textsc{APPLgrid} and \textsc{FastNLO} formats. The \textsc{PineAPPL}
library is interfaced to \textsc{MadGraph5\_aMC@NLO} (\textsc{mg5\_aMC}
henceforth), with which it has been developed and tested. In this respect,
\textsc{PineAPPL} supersedes
\textsc{APPLgrid}+\textsc{aMCfast}~\cite{Bertone:2014zva}.
However, \textsc{PineAPPL} can also be easily interfaced to any Monte Carlo generator,
e.g.\ \textsc{SHERPA}~\cite{Biedermann:2017yoi}, where it would complement
\textsc{MCgrid}~\cite{DelDebbio:2013kxa,Bothmann:2015dba}.

>>>>

In this paper we address the first of these two problems. Specifically, 
we develop \textsc{PineAPPL}, a library that allows any user to generate
fast-interpolation grids, accurate to any fixed order in the QCD and
EW couplings. The library supports variations of the factorisation
and renormalisation scales, and can be extended to include
resummation, and matching with a photon- and/or parton-shower.
The grids in the new \textsc{PineAPPL} format complement those
(accurate to fixed order in the strong coupling only) that can be generated
in the \textsc{APPLgrid} and \textsc{FastNLO} formats. The \textsc{PineAPPL}
library is interfaced to \textsc{MadGraph5\_aMC@NLO} (\textsc{mg5\_aMC}
henceforth), with which it has been developed and tested,
however it can be easily used with any Monte Carlo generator, 
e.g.\ \textsc{SHERPA}~\cite{Biedermann:2017yoi}. In this respect,
\textsc{PineAPPL} extends to EW corrections the scope of 
\textsc{aMCfast}~\cite{Bertone:2014zva} and
\textsc{MCgrid}~\cite{DelDebbio:2013kxa,Bothmann:2015dba}.

### 2020-09-03 paper/sec-pineappl.tex
Fix last-minute changes
Addressed Valerio's complaint

\item \textsc{PineAPPL} has been explicitly interfaced to \textsc{mg5\_aMC} (v3+)~\cite{Alwall:2014hca,Frederix:2018nkq},\footnote{A version of \textsc{mg5\_aMC} interfaced to \textsc{PineAPPL} can be downloaded from \url{https://launchpad.net/mg5amc-pineappl/trunk}, and it will be included as standard in the next \textsc{mg5\_aMC} release.} for which it supersedes \textsc{aMCfast}~\cite{Bertone:2014zva}; see appendix~\ref{app:sample-runcard} for an example of a runcard. While the usage of
\textsc{mg5\_aMC}(v3+) + \textsc{PineAPPL} is similar to that of 
\textsc{mg5\_aMC}(v2) + \textsc{aMCfast} + \textsc{APPLgrid}, it is however more
efficient. For instance, the usage of the latter typically required a substantial amount of memory (close to \SI{120}{\giga\byte} for Drell--Yan), if a grid for a differential distribution with more than 10 bins was generated.
The memory usage is substantially reduced (roughly \SI{1}{\giga\byte}) after \enquote{optimisation} of the grids, i.e.\ an optimisation of the grid representation in memory.
However, this required a two-step procedure: first, to produce an unoptimised grid in order to identify those parts where the cross section is either zero or extremely suppressed; second, after those parts are removed, to fill an optimised grid with a small number of bins.
\textsc{PineAPPL} avoids this by using a more space-efficient representation from the start. This leads to substantially faster run times, in particular for simple processes, because the grids do not need to be optimised 
and their combination is faster.

@@

\paragraph{Lagrange-interpolation grid.}
A different strategy is to partition a subset $H$ of the $(x_1, x_2, Q^2)$ space,
\begin{equation}
H = [x_\mathrm{min},x_\mathrm{max}]^2 \times [Q^2_\mathrm{min}, Q^2_\mathrm{max}] \ni (x_1, x_2, Q^2)
\end{equation}
along each axis into a small number of bins and to insert the phase-space weights $w$ into the corresponding discrete bin.
Using the bin centres and their values one already has a straightforward representation of eq.~\eqref{eq:weight-map}, but given a finite number of bins this approach usually yields an insufficient approximation for the cross section.
Increasing the number of bins improves the precision, but it also increases the space requirements.
This problem in turn in solved using interpolation methods, which increase the precision using the same number of bins.

>>>>

\item \textsc{PineAPPL} has been explicitly interfaced to \textsc{mg5\_aMC} (v3+)~\cite{Alwall:2014hca,Frederix:2018nkq},\footnote{A version of \textsc{mg5\_aMC} interfaced to \textsc{PineAPPL} can be downloaded from \url{https://launchpad.net/mg5amc-pineappl/trunk}, and it will be included as standard in the next \textsc{mg5\_aMC} release.} see appendix~\ref{app:sample-runcard} for an example of a runcard. While the usage of
\textsc{mg5\_aMC}(v3+) + \textsc{PineAPPL} is similar to that of 
\textsc{mg5\_aMC}(v2) + \textsc{aMCfast} + \textsc{APPLgrid}, it is however more
efficient. For instance, the usage of the latter typically required a substantial amount of memory (close to \SI{120}{\giga\byte} for Drell--Yan), if a grid for a differential distribution with more than 10 bins was generated.
The memory usage is substantially reduced (roughly \SI{1}{\giga\byte}) after \enquote{optimisation} of the grids, i.e.\ an optimisation of the grid representation in memory.
However, this required a two-step procedure: first, to produce an unoptimised grid in order to identify those parts where the cross section is either zero or extremely suppressed; second, after those parts are removed, to fill an optimised grid with a small number of bins.
\textsc{PineAPPL} avoids this by using a more space-efficient representation from the start. This leads to substantially faster run times, in particular for simple processes, because the grids do not need to be optimised 
and their combination is faster.

@@

\paragraph{Lagrange-interpolation grid.}
A different strategy is to partition a subset $H$ of the $(x_1, x_2, Q^2)$ space,
\begin{equation}
H = [x_\mathrm{min},x_\mathrm{max}]^2 \times [Q^2_\mathrm{min}, Q^2_\mathrm{max}] \ni (x_1, x_2, Q^2)
\end{equation}
along each axis into a small number of bins and to insert the phase-space weights $w$ into the corresponding discrete bin.
Using the bin centres and their values one already has a straightforward representation of eq.~\eqref{eq:weight-map}, but given a finite number of bins this approach usually yields an insufficient approximation for the cross section.
Increasing the number of bins improves the precision, but it also increases the space requirements.
This problem in turn is solved using interpolation methods, which increase the precision using the same number of bins.
